Power and Pitfalls of Generic Smart Contracts

### 2021-06-28 030_generics_attack.tex
Improved sentences

It is important that the \emph{Consistency of Shareholders} property holds
for \<TendermintValidators>, since
the shareholders of a \<TendermintValidators>
must be instances of \<TendermintED25519Validator> (as declared in the generic
signature of \<TendermintValidators>): class \<TendermintED25519Validator> enforces a
match between its public key, that identifies who can spend the rewards sent to the validator,
and its Tendermint identifier, that identifies which node of the blockchain
must do the validation work (Fig.~\ref{fig:validator}).
If it were possible to add a shareholder of another
type \<Attacker>, the latter could declare the identifier of a node that does not correspond to its
public key (see Fig.~\ref{fig:attacker}):
Tendermint would expect the node (belonging to the \emph{victim}) to do
the validation work while the owner
of the private key of the \<Attacker> could just wait for accrued awards to spend.
A sort of validator's slavery.
Sec.~\ref{sec:shared_entities} asserted that the \emph{Consistency of Shareholders}
property holds, at source level.
Namely, an \<attacker> (of type \<Attacker>) can only become shareholder
by accepting an ongoing sale \<offer> of shares through a call to
\<tv.accept(offer.cost, attacker, offer)> (see Fig.~\ref{fig:simple_shared_entity}).
This is impossible at source level, where that call does \emph{not} compile, since \<attacker> has type \<Attacker> that
is not an instance of \<S>, which has been set to \<TendermintED25519Validator>.
But a Hotmoka blockchain contains only the bytecode of \<SimpleSharedEntity>,
where the signature of \<accept> has been erased into
\<accept(BigInteger amount, PayableContract buyer, Offer offer)>.
Hence a blockchain transaction that invokes \<tv.accept(offer.cost, attacker, offer)>
\emph{does} succeed, since \<attacker> is an externally owned account and all such accounts
are \<PayableContract>s.
That transaction adds \<attacker> to the shareholders of \<tv>,
therefore violating the \emph{Consistency of Shareholders} property and allowing
validator's slavery.

>>>>

It is important that the \emph{Consistency of Shareholders} property holds
for \<TendermintValidators>: its shareholders
must be \<TendermintED25519Validator>s (as declared in the generic
signature of \<TendermintValidators>) that enforce a
match between their public key, that identifies who can spend the rewards sent to the validator,
and their Tendermint identifier, that identifies which node of the blockchain
must do the validation work (Fig.~\ref{fig:validator}).
If it were possible to add a shareholder of another
type \<Attacker>, the latter could declare the identifier of a node that does not correspond to its
public key (see Fig.~\ref{fig:attacker}):
Tendermint would expect the node (belonging to the \emph{victim}) to do
the validation work while the owner
of the private key of the \<Attacker> could just wait for accrued awards to spend.
A sort of validator's slavery.
Sec.~\ref{sec:shared_entities} asserted that the \emph{Consistency of Shareholders}
property holds, at source level.
Namely, an \<attacker> (of type \<Attacker>) can only become shareholder
by accepting an ongoing sale \<offer> of shares through a call to
\<tv.accept(offer.cost, attacker, offer)> (see Fig.~\ref{fig:simple_shared_entity}).
This is impossible at source level, where that call does \emph{not} compile, since \<attacker> has type \<Attacker> that
is not an instance of \<S>, which has been set to \<TendermintED25519Validator>.
But a Hotmoka blockchain contains only the bytecode of \<SimpleSharedEntity>,
where the signature of \<accept> has been erased into
\<accept(BigInteger amount, PayableContract buyer, Offer offer)>.
Hence a blockchain transaction that invokes \<tv.accept(offer.cost, attacker, offer)>
\emph{does} succeed, since \<attacker> is an externally owned account and all such accounts
are \<PayableContract>s.
That transaction adds \<attacker> to the shareholders of \<tv>,
therefore violating the \emph{Consistency of Shareholders} property and allowing
validator's slavery.
